Galois Slicing as Differentiation

### 2025-07-11 main/introduction.tex
Tweak intro example test to better foreshadow 3.3.

\begin{example}
  \label{ex:introduction-example}
  The following program is written in Haskell \cite{haskell} syntax, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then computing the sum of the numbers:
  \begin{displaymath}
    \begin{array}{l}
      \mathrm{query} :: \mathrm{Label} \to [(\mathrm{Label}, \mathrm{Int})] \to \mathrm{Int} \\
      \mathrm{query}\,l\,\mathit{db} = \mathrm{sum}\,[ n \mid (l',n) \leftarrow \mathit{db}, l \equiv l' ]
    \end{array}
  \end{displaymath}
  With $\mathit{db} = [(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, 1)]$, we will have $\mathit{query}\,\mathsf{a}\,\mathit{db}$ and $\mathit{query}\,\mathsf{b}\,\mathit{db}$ both evaluating to $1$.

@@

      % links
      \draw (top) -- (ioi);
      \draw (top) -- (oii);
      \draw (top) -- (iio);
      \draw (ioi) -- (ooi);
      \draw (ioi) -- (ioo);
      \draw (oii) -- (ooi);
      \draw (oii) -- (oio);
      \draw (iio) -- (ioo);
      \draw (iio) -- (oio);
      \draw (ooi) -- (bot);
      \draw (oio) -- (bot);
      \draw (ioo) -- (bot);
    \end{tikzpicture}
  \end{center}
  In both runs of the program, the output approximation lattice looks like this, where $1$ is the actual data point that was returned, and $\bot$ indicates that we are approximating this piece of data away:
  \begin{center}
    \begin{tikzpicture}[node distance=0.75cm]
      \node (top) at (0,0) {$1$};
      \node [below of=top] (bot) {$\bot$};
      \draw (top) -- (bot);
    \end{tikzpicture}
  \end{center}
  These are not the only choices of approximation lattices that we could have made. For the input, we have chosen a lattice that allows us to ``forget'' (approximate away) numbers in the input, but not any of the other data or the overall structure. However, other choices are also useful. Indeed, one of the aims of this work is to clarify how to choose an approximation structure appropriate for different tasks by use of type information. We elaborate on this further in \secref{models-of-total-gps}.

>>>>

\begin{example}
  \label{ex:introduction-example}
  The following program is written in Haskell syntax \cite{haskell}, using a list comprehension to filter a list of pairs of labels and numbers to those numbers with a given label, and then computing the sum of the numbers:
  \begin{displaymath}
    \begin{array}{l}
      \mathrm{query} :: \mathrm{Label} \to [(\mathrm{Label}, \mathrm{Int})] \to \mathrm{Int} \\
      \mathrm{query}\,l\,\mathit{db} = \mathrm{sum}\,[ n \mid (l',n) \leftarrow \mathit{db}, l \equiv l' ]
    \end{array}
  \end{displaymath}
  With $\mathit{db} = [(\mathsf{a}, 0), (\mathsf{b}, 1), (\mathsf{a}, 1)]$, we will have $\mathit{query}\,\mathsf{a}\,\mathit{db}$ and $\mathit{query}\,\mathsf{b}\,\mathit{db}$ both evaluating to $1$.

@@

      % links
      \draw (top) -- (ioi);
      \draw (top) -- (oii);
      \draw (top) -- (iio);
      \draw (ioi) -- (ooi);
      \draw (ioi) -- (ioo);
      \draw (oii) -- (ooi);
      \draw (oii) -- (oio);
      \draw (iio) -- (ioo);
      \draw (iio) -- (oio);
      \draw (ooi) -- (bot);
      \draw (oio) -- (bot);
      \draw (ioo) -- (bot);
    \end{tikzpicture}
  \end{center}
  In both runs of the program, the output approximation lattice looks like this, where $1$ is the actual data point that was returned, and $\bot$ indicates that we are approximating this piece of data away:
  \begin{center}
    \begin{tikzpicture}[node distance=0.75cm]
      \node (top) at (0,0) {$1$};
      \node [below of=top] (bot) {$\bot$};
      \draw (top) -- (bot);
    \end{tikzpicture}
  \end{center}
  These are not the only choices of approximation lattices that we could have made. For the input, we have chosen a lattice that allows us to ``forget'' (approximate away) numbers in the input, but not the labels or the structure of the list itself. However, other choices are also useful. Indeed, one of the aims of this work is to clarify how to choose an approximation structure appropriate for different tasks by use of type information. We elaborate on this further in \secref{semantic-gps}.
